---
title: "Life Satisfaction Analysis between Full-time and Non-Full-time Workers"
author: "Wen Yu, Ani Hovhannisyan, Yue He(Joy)"
date: "2025-04-21"
output: html_document
---

## Introduction & Methods
This report analyzes the status of life satisfaction among people working full-time and non-full-time in Germany.
We test the following Null-Hypothesis:
H0: No difference in life satisfaction between Full-time and Non-full-time workers in Germany

We used the variable `mainact` to define employment type.
It was recoded into a binary variable `work_type`, where “Paid work” was labeled as “Full-time” and all other types (e.g., Retired, Education, Housework.etc) were grouped as “Not full time”.

```{r}

library(haven)
library(dplyr)
ess11_de <- read_sav('ESS11-subset.sav') #Please put the data file I uploaded in the whatsapp chat in your 'Sandbox' so that you can import the original data
ess11_clean <- ess11_de %>% 
  select(mainact, stflife) %>% 
  mutate(stflife=ifelse(stflife==88, NA, stflife))%>%#set the 88 into NA, so that we can delect them in the following code
  filter(!is.na(mainact), !is.na(stflife)) %>% #select the variables that we need and get rid of the empty values
  mutate(mainact=as_factor(mainact)) #convert the mainact into the proper categorical variables with readable value in R
table(ess11_clean$mainact)#check if the data has been cleand successfully
ess11_analysis <- ess11_clean %>%
  mutate(work_type=ifelse(mainact=='Paid work',
                          'Full-time',
                          'Not full time')) #divide the work_type into two categories: 'Full-time' and 'Not full time' and add an column to the data


```


## Regression Results
According to the linear regression results, the Mean of life satisfaction score for full-time workers was 7.66, while for non-full-time workers it was 7.84. There was no statistically significant difference in life satisfaction between full-time and part-time workers (β = 0.18, p = 0.274 > 0.05). Meanwhile, the 95% CI is between (-0.14,0.50), includes 0. Suggesting that employment type may not be a strong predictor of life satisfaction in this sample. We can't reject the Null-Hypothesis.

```{r}

ess11_analysis %>%
  group_by(work_type)%>%
  summarise(main_satisfaction=mean(stflife))
#by the code above, we can see that the mean of main_satisfaction of Full-time is less than that of Not full time
model <- lm(stflife ~ work_type, data = ess11_analysis, na.action=na.omit)
summary(model)#run the regression model, I will send the analysis of this in the whatsapp chat

```

## Boxplot
```{r}

library(ggplot2)
library(tidyverse)
# draw a boxplot
ggplot(ess11_analysis, 
       aes(x = work_type,
           y = stflife,
           fill = work_type)) +
  geom_boxplot() +
  labs(
    title = "Life Satisfaction by Employment Type",
    x = "Employment Type",
    y = "Life Satisfaction Score (0-10)",
    fill = "Work Type"
  )

# dir.create("data_clean", showWarnings = FALSE)
# write_rds(ess11_analysis, "data_clean/ess11_analysis.rds")

```

## Conclusion
Our analysis does not support the hypothesis that there's no significant difference in life satisfaction between Full-time and Non-full-time workers in Germany.
However, The sample size (N=440) may not be sufficient to detect small effects. For further analysis, it may be necessary to control for latent variables (age, gender, income, etc.), and cross-test for them.

## Members & Contributions
Wen Yu: Data Import, Cleaning & Regression Analysis  
Yue He(Joy): Visualization (Boxplot), Report Writing  
Ani Hovhannisyan: GitHub Repository Setup, README, Integration
